Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-08-19 templates_aimusic2020/paper.tex
Updates from Overleaf

\begin{abstract}
Digital sound artists often require a variety of percussive samples for their music. For more than two decades research involving digital synthesis, heuristic search, and neural networks has been used for the generation of novel sounds. Our goal is to generate one-shot percussive samples by leveraging modern AI technologies alongside scalable signal generation methods. We centered our approach around the combination of two central components: \begin {enumerate*} [label=(\roman*)] \item a \emph{virtual ear} capable of evaluating the proximity of unheard sounds to  and \item a dynamic \emph{virtual synthesizer} with a rich set of tractable parameters\end{enumerate*}. We present a generative pipeline that utilizes robust digital signal processing methods guided by supervised learning towards the generation of one-shot percussive sounds. We present our findings and measurements of the various approaches taken. Advantages, shortcomings and practicality of the employed methodologies are demonstrated. We share our curated dataset of sounds and codebase\footnote{\url{https://github.com/imilas/Synths_Stacks_Search}} which can be used for its further expansion.
\keywords{Automatic Synthesizer design. Machine Listening. Sound Analysis. Novelty and Originality}
\end{abstract}

\section{Preamble}
\subsection{Drums and Percussion}
Drums are a subcategory of percussive instruments. Drum sounds such as kick and snare are two of the most universally recognized examples of percussion~\cite{barry2005drum}. Almost any object which is struck to produce noise can be referred to as a drum~\cite{latham2002oxford}. The term \enquote{Percussion} often encompasses all drums and may include the additional instruments which may be rubbed, plucked, blown, etc; A more strict definition of percussion can be used interchangeably with drums~\cite{latham2002oxford}. We adapt this strict definition for this work as our concern is not the delineation between the two overlapping concepts, but with extraction and creation of novel sounds with characteristics shared between the two. As a result, we often use the terms \enquote{drum} and \enquote{percussion} interchangeably.

\subsection{Motivation and Goals}
Digital recordings of novel, one-shot\footnote{A single hit on the drum that captures its capabilities} drum sounds are not easy or cheap to find. Our goal here is to create a new, virtual source of novel drum sounds, which could then be used in electronic music compositions.

By relying on recordings of \textit{real life} drum sounds, we are limited by what instruments exist in the real world and whether or not we have access to clean, one-shot recordings. Our hypothesis is that the virtualization of sound generation can alleviate these material limitations.

 \begin{figure}[t!]
    \begin{center}
    \textbf{Pipeline Design}
    \makebox[\textwidth]{
    \fbox{\includegraphics[width=1.1\linewidth]{images/pipeline.pdf}}}
    \end{center}
    \caption{A blueprint of our desired pipeline which allows each component to be implemented in a number of ways. Our implementations of this pipeline allow for easy parallelization when needed. 
    }
\label{fig:pipeline_outline}
\end{figure}

\subsection{Related Works}
\label{related_works}
\begin{center}
\begin{table}[h]
\resizebox{\linewidth}{!}{\begin{tabular}{|c c c c|} 
\hline
Work & Feature extraction & Synthesis & Specilization & \hline
Oord 2017 & CNN & CNN &Speech& \hline
Yee-King 2018 & LSTM on Paremters & DSP & Synth Pads & \hline
Aouameur 2019 & Latent layer& Decoding of Latent Layers & Percussion & \hline
Ramires 2020 & Latent layer & FeedForward Network & Percussion & \hline
Yamamoto 2020 & GAN & GAN & Speech & \hline
\end{tabular}}
\caption{Quick reference of recent related works}
\end{table}
\label{table:recent_works}
\end{center}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel}. Most relevant to us are recent works by Aoumaeur et al. in which variational AutoEncoders (VAE's) have been utilized for generation of percussive samples~\cite{aouameur2019neural}; As well as Ramires et al's work in generation of percussive sounds by guiding the output of a feedforward neural network via a small set of latent features~\cite{ramires2020neural}. 
Automatic programming of virtual synthesizers has long been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines~\cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al.~\cite{yee2018automatic} used Long short-Term Memory (LSTM) models and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain even with random search. Since this work appears more focused on pads and textures rather than drums, feature matching appears to not be concerned with the envelope of the sounds but rather the frequency content within arbitrary time windows. Yet another recent work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio~\cite{esling2019universal}. As stated before, our work explores the rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation.

\subsection{Our Approach}
To virtually create novel drum sounds, we need a tractable, generative source of audio; One that deterministically produces audio based on a set of instructions. Additionally, we need a method of evaluation to help us determine which sounds resemble drums and are worth keeping. To enable heuristic search, we would like to know what instructions caused our audio source to make the sounds we liked. Given these two components capable of \textit{generation} and \textit{evaluation} of sounds, we build a generative pipeline of audio as depicted in figure~\ref{fig:pipeline_outline}. The success of this pipeline is dependent on the characteristics of the components responsible for the generation and evaluation of the sounds. We call these two components the \emph{virtual synthesizer} and \emph{virtual ear} respectively.  
\begin{figure}[t!]
    \begin{center}
    \textbf{Data Overview}
    \makebox[\textwidth]{
    \includegraphics[width=1.1\linewidth]{images/venn_data.pdf}}
    \end{center}
    \caption{ 
    Scales are symbolic (True scales are not measurable and/or subjective). Our positive samples, $\mathcal{T^{+}}$, is a small fraction of a wide variety of percussive sounds that are conceivable. The size of $\mathcal{T^{-}}$ can be set to any number(we can generate any number of random samples), but we prefer to keep it no larger than  $\mathcal{T^{+}}$. $\mathcal{H}$ is a series of sounds sent to the ear for classification. The ideal ear will detect  $\mathcal{H^{+}}$, the overlap of $\mathcal{H}$ and $\mathcal{N}$, with high precision and recall. 
    }
\label{fig:ven_data}
\end{figure}

\subsection{Virtual Synthesizer Design}
Sound is product of physical disturbances, causing vibrations in our mediums. Vibrations are perpetuated through air as part of an expanding, spherical front ~\cite{cook1999chap4}. A sound wave can be viewed as the result of a function which governs amplitude through time, where time and amplitude exist in continuous dimensions. Waves can be approximated via a series of samples, associating time steps to a discrete range of amplitude values.~\textit{Digital synthesis} of audio is the process of creating these discrete values.
Unlike the majority of the works mentioned in Section~\ref{related_works}, we do not employ probabilistic models for the synthesizer component. Our decision is based on the following factors:
\begin{itemize}
    \item \textit{Interpretability and Expansion}:
    \item \textit{Novelty and Creativity}: Given large sets of examples, generative models such as VAEs and GANS have proven capable of generating samples similar to the exemplars. we are not aiming for perfect imitations, rather, we want to work with the limitations of any tractable sound source to create its novel approximations of a given sound category.     
    \item \textit{Efficiency}:

\end{itemize}

\begin{figure*}[htbp]
\label{fig_example_sine}
\centering
\includegraphics[width=0.45\linewidth,angle =-90 ]{images/periodic_function.png}
\caption{A computer can simulate waveforms by utilizing periodic functions. Digital waveforms are discrete approximations of analogue waves. \cite{mitchell2009basicsynthChap5} }
\end{figure*}

@@

\subsection{How Can a Computer Evaluate Sounds?}
Automatic evaluation of sounds is an essential component of our work: A thorough manual evaluation of outputs is not possible when hundreds can be created in a second. In our work, we find simple features such as frequency content (high pitch vs low pitch), length, and envelope (change in loudness, how fast the sounds reaches its maximum loudness and fades away) to be powerful for the categorization of drum sounds. Later on we will discuss our algorithms for extraction of these and other features in~\ref{implementation}. We will also discuss how these extracted features were used to train models that can automatically categorize new sounds.

\section{Implementation}
We conceptualized a general pipeline for approximation of sounds: A synthesizer creates random noise while a virtual ear rapidly evaluates the outputs and outputs a score which is used for the separation of undesired outputs from desired ones. This approach assumes that a fraction of the randomly generated sounds can be substituted for percussion and that the synthetic ear will assign higher evaluation scores to this subset. 
\begin{itemize}
    \item \textit{Virtual Synthesizer}: A flexible, deter\-min\-istic, and tract\-able gener\-ator which can create audio. 
    \item \textit{Virtual Ear}: An ear that returns an evaluation of an audio sample; estimating the effectiveness of an audio sample's fulfillment of a producers requirements. The ear's evaluation guides the generation process towards a desired path, making it a crucial component of our pipeline. 
\end{itemize}
We wish to leverage AI technologies, heuristic search, and DSP methods for the synthesis of novel audio. The work here is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. We demonstrate our work in combining machine listening and automatic synthesizer programming to find pieces of audio which can be used as drums. Using the generation of short, percussive audio samples as a starting point, this project is a proof of concept for a modular, generative pipeline of novel, one-shot audio samples. Design decisions were made such that this approach can be modified for creation of other, non-percussive sound-types. 
Instead of learning weights and parameters in an audio-generation neural network, we wish to generate, search, and tune synthesizers to produce percussion sounds.

@@

\end{document}

>>>>

\begin{abstract}
Digital sound artists often require a variety of percussive samples for their music. For more than two decades research involving digital synthesis, heuristic search, and neural networks has been used for the generation of novel sounds. This work is a proof of concept in automatic programming of synthesizers towards their best approximation of drum sounds. We centered our approach around the combination of two central components: \begin {enumerate*} [label=(\roman*)] \item a \emph{virtual ear} capable of evaluating the proximity of unheard sounds to  and \item a dynamic \emph{virtual synthesizer} with a rich set of tractable parameters\end{enumerate*}. We present a generative pipeline that utilizes classical digital signal processing functions guided by supervised learning towards the generation of one-shot percussive sounds. We present our findings and measurements of a number of approaches taken. Advantages, shortcomings and practicality of the employed methodologies are demonstrated. We share our curated dataset of sounds and codebase\footnote{\url{https://github.com/imilas/Synths_Stacks_Search}} which can be used for its further expansion.
\keywords{Automatic Synthesizer design. Machine Listening. Sound Analysis. Novelty and Originality}
\end{abstract}

\section{Preamble}
\subsection{Drums and Percussion}
Drums are a subcategory of percussive instruments. Almost any object which is struck to produce noise can be referred to as a drum~\cite{latham2002oxford}. Kicks and snares are two of the most universally recognized examples of percussion~\cite{barry2005drum}. The term \enquote{Percussion} often encompasses all drums and may include the additional instruments which may be rubbed, plucked, blown, etc; A more strict definition of percussion can be used interchangeably with drums~\cite{latham2002oxford}. We adapt this strict definition for this work as our concern is not the delineation between the two overlapping concepts, but with extraction and creation of novel sounds with characteristics shared between the two. As a result, we often use the terms \enquote{drum} and \enquote{percussion} interchangeably.

\subsection{Motivation and Goals}
Digital recordings of novel, one-shot\footnote{A single hit on the drum that captures its capabilities} drum sounds are not easy or cheap to find. Our goal here is to create a new, virtual source of novel drum sounds, which could then be used in electronic music compositions. By relying on recordings of \textit{real life} drum sounds, we are limited by what instruments exist in the real world and whether or not we have access to clean, one-shot recordings. Our hypothesis is that the virtualization of sound generation can alleviate these material limitations.

 \begin{figure}[t!]
    \begin{center}
    \textbf{Pipeline Design}
    \makebox[\textwidth]{
    \fbox{\includegraphics[width=1\linewidth]{images/pipeline.pdf}}}
    \end{center}
    \caption{A blueprint of our desired pipeline which allows each component to be implemented in a number of ways. Our implementations of this pipeline allow for easy parallelization when needed. 
    }
\label{fig:pipeline_outline}
\end{figure}

\subsection{Related Works}
\label{related_works}
\begin{center}
\begin{table}[h]
\resizebox{\linewidth}{!}{\begin{tabular}{|c c c c|} 
\hline
Work & Feature extraction & Synthesis & Specilization & \hline
Oord 2017 & CNN & CNN &Speech& \hline
Yee-King 2018 & LSTM on Paremters & DSP & Synth Pads & \hline
Aouameur 2019 & Latent layer& Decoding of Latent Layers & Percussion & \hline
Ramires 2020 & Latent layer & FeedForward Network & Percussion & \hline
Yamamoto 2020 & GAN & GAN & Speech & \hline
\end{tabular}}
\caption{Quick reference of recent related works}
\end{table}
\label{table:recent_works}
\end{center}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel}. Particularly relevant are the utilization of variational autoencoders (VAE's) for generation of percussive samples~\cite{aouameur2019neural} and generation of percussive sounds by decoding a small set of latent features~\cite{ramires2020neural}.
Automatic programming of virtual synthesizers has long been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines~\cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al.~\cite{yee2018automatic} used Long short-Term Memory (LSTM) models and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain even with random search. Since this work appears more focused on pads and textures rather than drums, feature matching appears to not be concerned with the envelope of the sounds but rather the frequency content within arbitrary time windows. Yet another recent work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio~\cite{esling2019universal}. As stated before, our work explores the rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation.

\subsection{Our Approach}
To virtually create novel drum sounds, we need a tractable, generative source of audio; One that deterministically produces audio based on a set of instructions. Additionally, we need a method of evaluation to help us determine which sounds resemble drums and are worth keeping. To enable heuristic search, we would like to know what instructions caused our audio source to make the sounds we liked. Given these two components capable of \textit{generation} and \textit{evaluation} of sounds, we build a generative pipeline of audio as depicted in figure~\ref{fig:pipeline_outline}. The success of this pipeline is dependent on the characteristics of the components responsible for the generation and evaluation of the sounds. We call these two components the \emph{virtual synthesizer} and \emph{virtual ear} respectively.

\section{Virtual Synthesizer Design}
\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |c|c|c| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & tone type\\
IsNoise & boolean & whether to \newline use OSC type to generate noise\\
Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start$<$1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cuttoff & 0-20000hz & -\\
LP filter Cuttoff & 20000-HP & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer submodule Parameters. Despite the simplicity of the parameters and our efforts at constraining the ranges, the number of parameters that can be randomly chosen for each submodule is in the order of $10^{15}$ }
\label{table:submodule_params}
\end{table}

Sound is product of physical disturbances, causing vibrations in our mediums. Vibrations are perpetuated through air as part of an expanding, spherical front ~\cite{cook1999chap4}. A sound wave can be viewed as the result of a function which governs amplitude through time, where time and amplitude exist in continuous dimensions. Waves can be approximated via a series of samples, associating time steps to a discrete range of amplitude values.~\textit{Digital synthesis} of audio is the process of creating these discrete values.
Unlike the majority of the works mentioned in Section~\ref{related_works}, we do not employ probabilistic models for the synthesizer component. Our decision is based on the following factors:
\begin{itemize}
    \item \textit{Novelty and Creativity}: Given large sets of examples, generative models such as VAEs and GANS have proven capable of generating samples similar to the exemplars. In this work, we are not aiming for perfect imitations; Following Boden's definition of exploratory creativity as an emergent property of generative work within confined rule sets~\cite{boden2009computer}, we work within the limitations of any tractable sound source to create its  approximations of a given sound category. Within the music realm, prominent instance of exploratory creativity is the persistent aesthetics of 8-bit soundtracks, remaining popular long after the original constraints are lifted~\cite{collins2007loop}.
    \item \textit{Interpretability}: Neural networks are often described as black boxes~\cite{basheer2000artificial}. Their highly recursive structure makes modern explanation methods such as saliency maps unreliable~\cite{rudin2019stop}.  
    \item \textit{Speed of Rendering at high sampling rates}: Despite the utilization of powerful GPUs, the standard sampling rate in most audio generation work utilizing neural is under 24 khz \cite{yamamoto2020parallel,oord2017parallel,aouameur2019neural,ramires2020neural}. However, a significant number of untrained human ears can detect a change in quality of audio between sampling rates of 192 khz and the industry standard of 44.1 khz \cite{reiss2016meta}. In this work, we fix our sampling rate to the 48 khz standard. 
\end{itemize}

\begin{figure*}[htbp]
\label{fig_example_sine}
\centering
\includegraphics[width=0.45\linewidth,angle =-90 ]{images/periodic_function.png}
\caption{A computer can simulate waveforms by utilizing periodic functions. Digital waveforms are discrete approximations of analogue waves~\cite{mitchell2009basicsynthChap5} }
\end{figure*}

@@

We opted for a set of classical DSP methods to build our synthesizer. For our project, we used the python based Pippi library for sound generation\footnote{https://github.com/luvsound/pippi}. This library uses a C back-end\footnote{https://github.com/PaulBatchelor/Soundpipe} and focuses on fast offline generation of audio signals. Our virtual synthesizer contains a set of one or more submodules. Each submodule is a self-contained noise making unit. Submodules have identical sets of parameters, but widely different outputs can be achieved depending on the values assigned. The set of parameters available to each submodule is highlighted in Table~\ref{table:submodule_params}. The sonic output of the virtual synthesizer is the normalized addition of the sonic output of its submodules. Our implementation of a synthesizer can have any number of submodules. The parameters that dictate the output signal of each submodule as well as the range of values each parameter can take are shown in table~\ref{table:submodule_params}. We call the number of submodules in each virtual synthesizer the \textit{stack size}. We call the sets of parameter values that characterize a synthesizer's submodules a \textit{program} (analogous to a preset for a VST).

\section{Virtual Ear}

We are not looking for the perfect imitation of organic drums using a synthesizer. We seek to imitate drums using a synthesizer and even \textbf{prefer} for its generations to retain novel, unusual characteristics. The task assigned to the virtual ear is the rare acceptance of drum-like sounds and the inevitable rejection of most \enquote{noise} outputs from the virtual synthesizer. By definition, we cannot anticipate what these novel drums will sound like. Without manual inspection, all we know is that the accepted sounds are synthesizer noises which the synthetic ear has determined to be similar to drum-sounds. The virtual ear makes two critical decisions: 
\emph{
\begin{quote}
\text{Decision.1 Could the sound be used as a drum?}\label{Decision.1}
\\
\text{Decision.2 If it does sound like a drum, what type of drum should it be?}\label{Decision.2}
\end{quote}
 }

\begin{figure}[htbp!]
    \begin{center}
    \textbf{Data Overview}
    \makebox[\textwidth]{
    \includegraphics[width=0.95\linewidth]{images/venn_data.pdf}}
    \end{center}
    \caption{ 
    Scales are symbolic (true scales are not measurable and/or subjective).  $\mathcal{N}$ is the set of percussive sounds a synthesizer is capable of making. Our positive samples, $\mathcal{T^{+}}$, is a small fraction of a wide variety of percussive sounds that are conceivable. The size of $\mathcal{T^{-}}$ can be set to any number(we can generate as any number of random samples), but we prefer to keep it no larger than  $\mathcal{T^{+}}$. $\mathcal{H}$ is a series of sounds sent to the ear for classification. The ideal ear will detect  $\mathcal{H^{+}}$, the overlap of $\mathcal{H}$ and $\mathcal{N}$, with high precision and recall. 
    }
\label{fig:ven_data}
\end{figure}
\hypref{Decision.1} requires knowledge of what drums \textbf{do not} sound like, or knowledge of an infinitely large set, which cannot be fully represented via examples. An important consideration is that the source of sounds used in training the model (organic drum sounds) will be fundamentally different from the source of unlabeled sounds we wish to categorize (noise from a synthesizer). This issue is reflective of the OSR problem~\cite{geng2020recent,mundt2019open}. Figure~\ref{fig:ven_data} highlights a number of caveats with this approach. we denote $\mathcal{N}$ as the set of percussive sounds a synthesizer is capable of making. $\mathcal{T^{+}}$, our positive examples, is a set of sound extracted from material drums. $\mathcal{T^{-}}$, is a small subset of an infinite set which is meant to represent all noises which the synthesizer is capable of making. It likely includes sounds in $\mathcal{N}$. could be used as drums. During the hearing test, the virtual ear continually receives synthesizer noises ($\mathcal{H}$) for classification. The performance of the ear is measured by its precision and recall in finding the overlap between $\mathcal{H}$ and $\mathcal{N}$. It is important to ensure that the change in learning domains---particularly with the case of positive examples of percussive instruments---does not interfere with transformation of knowledge from the training of the ear to the hearing test. To maximize the transference of knowledge from one domain to another, we seek to learn from agnostic feature sets that capture fundamental characteristics of the data points.
\subsection{Feature Extraction}
To train our virtual ear, we require generalizable, domain agnostic features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT) ~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used its reconstruction, perhaps it can be utilized as a source of fundamental features necessary for audible signal categorization. We defined 3 STFT transformation functions which we believe to capture important, unique attributes of percussive sounds (figure ~\ref{fig:stackspectrums}). Our transformations are GPU based, allowing their application in tandem to the training of the classifier or tuning of STFT transformation during hyper-paramter optimization (see Section~\ref{---}).

\subsection{Classifiers}
\label{classifiers}
Two groups of classifiers are implemented: \emph{two phased ears} (TPEs) and \emph{embedding ears} (EEs). TPEs are a combination different models specializing in ~\hypref{Decision.2} or ~\hypref{Decision.1} and trained only with our transformation functions. EEs are trained on a highly compressed, automatically encoded representation of sounds and can give simultaneous answers to both questions. These encodings are extracted from the latent layer of a percussion autoencoder.

@@

\begin{appendices}
\chapter{Datasets}
\label{appendix:A}
\begin{table}[h!]
\centering
\begin{tabular}[width=\paperwidth]{|l|l|l|l|l|l|l|l|l|l|}
\hline
DB Name & kick & snare & clap & tom\_high & tom\_mid & tom\_low & hihat\_closed &  hihat\_open & rim \\ \hline
MixedDB & 648 & 732 & 118 & 179 & 139 &  188 & 187 & 280 & 105 \\\hline
\end{tabular}
\caption{Database 1: Mixed sources}
\label{db:self}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
\hline
DB Name & kick & snare & clap & tom & clap & hat & rim & shaker  \\ \hline
RadarDB & 1054 & 842   & 353 & 349 &  353 & 1561& 131 & 121 \\ \hline
\end{tabular}
\caption{Database 2: Royalty free sounds sourced from \enquote{Music Radar}}
\label{db:radar}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
 DB Name & kick & snare & clap & hat & other \\\hline
 FreeDB & 533 & 372 & 230 & 105 & 281 \\ \hline
\end{tabular}
\caption{Database 3: Free sounds sourced from the \enquote{Sample Swap} project. Simplified for our purposes. The version available for download contains more sample groups. The \enquote{other} category contains a variety of percussive sounds.}
\label{db:sampleswap}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
 Synth Noise Type & 1 Stack & 3 Stacks  & 5 Stacks \\ \hline
 Number of Examples & 2000 & 2000 & 2000 \\ \hline
\end{tabular}
\caption{Database of random noise examples from our virtual synthesizers}
\label{db:noise}
\end{table}

%=============
\chapter{Audio Transformation functions}
\label{appendix:B}
\begin{figure}
\centering
\textbf{Visual Representation of Raw Features}\par\medskip
    \subcaptionbox{Recorded hat sample}{    \includegraphics[width=1\columnwidth]{images/ff1.pdf}
    }
    \subcaptionbox{Randomly generated audio with percussive qualities, resembling a tight snare}{\includegraphics[width=1\columnwidth]{images/ff2.pdf}}
    \subcaptionbox{A randomly generated noise with a percussive envelop but non-percussive frequency features (modulated pitch)}
    { \includegraphics[width=1\columnwidth]{images/ff3.pdf}}
\caption{Graphed representation of features extracted for 3 different samples. Sample $a$ is a recorded hat from our database. sample $b$ is an example of randomly generated noise with percussive qualities that we found suitably similar to a snare sound. Sample $c$ is an example of a randomly generated noise where the spectrum features are necessary for proper classification.}
\label{fig:stackspectrums}
\end{figure}
\begin{enumerate}
\item Envelope Transformation: Represents changes in loudness for the duration of the signal. Using STFT we generate a matrix $M_{i \times j}$ with rows $i$ and columns $j$ corresponding to time steps and frequency bins respectively. Values $v_{i \times j}$ indicating the magnitude of the frequency bin $j$ at each time-step $i$. Information about the envelope of the signal can be extracted by summing the values of $M$ for each time-step (or row $i$), giving us a feature vector that is normalized to the range of 0 to 1. The information contained in this vector is an alternative to Root-Mean-Square measurements of a sliding window over the signal.
\item Frequency Transformation: A static, normalized snap-shot of the the frequencies present within the audio. The calculation of this feature vector is similar to the envelope, but the summation is done along the frequency axis. Another important distinction is that since capturing an adequate frequency resolution is important for this transformation, we utilized shorter hop-sizes and wider windows. A Mel Scale transformation was also applied in hopes that the captured features better represent human perception of frequencies. 
\item Spectrum Transformation: This function is simply a Mel Scaled STFT with its values normalized from 0-1. Spectrograms contain more information about the original signal but their analysis requires more complex computational methods.
\end{enumerate}

%================
\chapter{Virtual Ear Classifier Definitions}
\label{appendix:C}
\section{Two Phased models}
Using the described features, we trained several neural network models for Phase 1 and 2 in the Pytorch environment. The task of Phase 1 is to separate drums from not-drums (DrumVsNotDrum, or DVN). The task of Phase 2 is to categorize drums and percussion (DrumVsDrum, or DVD). We kept our feature space small, making it viable for feature selection and model design to be done on a trial and error basis. For all models, accuracy is calculated by prediction of all test dataset labels and the loss function and optimizer are Categorical-CrossEntropy and Adam respectively. Training continues until no reduction in loss and accuracy is observed in 10 epochs.  All activation functions are PReLU:
\begin {enumerate}
\item FC-DVN: Fully connected network trained on Envelope features, reaching 97\% accuracy on our test data for Phase 1. With size of 10x5x10.
\item CNNLSTM-DVN: A combination of CNN and LSTM models, where the CNN model extracts higher level features that are fed temporally to an LSTM cell. This model is trained on spectrum data and reaches 98\% accuracy on our test set. Its structure is the combination of a CNN with 2 output channels and kernel size $(7,3)$; Followed by an LSTM model of hidden size 800 and a fully connected layer of size 20x2.
\item E+F-DVD: A fully connected model trained on a concatenation of envelope and frequency features. Reaching 80\% accuracy for 6-way drum categorization in Phase 2. Size of 50x10x2x6.
\item CNN-DVD: A CNN model trained on Spectrum features. Reaching 82\% accuracy in a 6-way drum categorization in Phase 2. A combination of a CNN model with output channel size of 4, kernel of size of 5, another CNN model with output channel size of 8 and kernel of size 3. Followed by a fully connected network of shape 100x20x6.
\item FC-DVD: Fully connected 3 layer neural net with 78\% accuracy for 6-way drum categorization in Phase 2. Size of 400x200x50.
\end{enumerate}
Unlike EE models, parameters are hand-picked and un-tuned. Higher accuracy rates in these models do not necessarily translate to higher agreeableness with surveyors. Since the positive sounds in the hearing test are sourced from a synthesizer, model accuracy on test data alone cannot be relied upon when the domain of sounds 
being categorized is switched.
\section{Embedding Models}

\begin{table}[[h]
\begin{tabular}{|p{28mm}|p{25mm}|p{23mm}|p{50mm}|}
\hline
Layer-\# & Out Shape & Param Num & Details  \\ \hline
Conv2d-1 & [-1, 8, 30, 20] &   208 & Encoder's input \newline
Num. Channels:8\newline
kernel:5x5\newline                  
stride:1\newline    
padding:2 \\ \hline
ReLU-2 & [-1, 8, 30, 20] &   0 & \\  \hline
MaxPool2d-3 & [-1, 8, 15, 10] & 0 &  kernel:5x5 \newline
stride:2 \\ \hline
Dropout-4 & [-1, 8, 15, 10] & 0 &  \\ \hline
Linear-5 & [-1, 8] & 9,608 & Encoder's output \\ \hline
Linear-6 & [-1, 256] & 2,304 & Decoder's Input \\ \hline
Dropout-7 & [-1, 256] & 0 &  \\ \hline
Linear-8 & [-1, 600 ] &  154,200& Decoder's output\\ \hline
\end{tabular}
\caption{CNN model design with latent size of 8. 30 and 20 are the assumed frequency bins and step size. Total number of parameters is 166,320. }
\label{table:CNNAUTOENCODER}
\end{table}

\begin{table}[[h]
\begin{tabular}{|p{28mm}|p{25mm}|p{23mm}|p{50mm}|}
\hline
Layer-\# & Out Shape & Param Num & Details  \\ \hline
Linear-1 & [-1, 128]  & 76,928 & Encoder's input \\ \hline
Dropout-2 & [-1, 128] & 0 &  \\ \hline
Linear-3 & [-1, 8] & 9,608 & Encoder's output \\ \hline
Linear-4 & [-1, 128] & 2,304 & Decoder's Input \\ \hline
Dropout-5 & [-1, 128]  & 0 &  \\ \hline
Linear-6  & [-1, 600 ] &  77,400 &Decoder's output\\ \hline
\end{tabular}
\caption{Fully connected model with only 1 hidden dimension for encoder and decoder. Designed assumes latent size of 8. 30 and 20 are the assumed frequency-bins and step-size values. Total number of parameters is 156,512.}
\label{table:FC1_AUTOENCODER}
\end{table}

\begin{table}[[h]

\begin{tabular}{|p{28mm}|p{25mm}|p{23mm}|p{50mm}|}
\hline
Layer-\# & Out Shape & Param Num & Details  \\ \hline
Linear-1 & [-1, 128]  & 76,928 & Encoder's input \\ \hline
Dropout-2 & [-1, 128] & 0 &  \\ \hline
Linear-3 & [-1, 32]  & 4,128 & \\ \hline
Dropout-4 & [-1, 128] & 0 &  \\ \hline
Linear-5 & [-1, 8] & 9,608 & Encoder's output \\ \hline
Linear-4 & [-1, 32] & 2,304 & Decoder's Input \\ \hline
Dropout-5 & [-1, 32]  & 0 &  \\ \hline
Linear-4 & [-1, 128] & 2,304 & \\ \hline
Dropout-5 & [-1, 128]  & 0 &  \\ \hline
Linear-6  & [-1, 600 ] &  77,400 &Decoder's output\\ \hline
\end{tabular}
\caption{Fully connected model with 2 hidden dimensions for encoder and decoder. Designed assumes latent size of 8. 30 and 20 are the assumed frequency-bins and step-size values. Total number of parameters is 163,232.}
\label{table:FC2_AUTOENCODER}
\end{table}

\end{appendices}
\end{document}
